Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface

### 2022-04-18 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Final bits before submission
Unnumber extra sections
Modifications for submission

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface"
author:
  - name: Brett E. Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Josh R. Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon L. Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation:  Woods Hole Oceanographic Institution,Plymouth State University
  - name: Sydney D. Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark L. Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark D. Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Environmental Science and Policy, 17 High St., Plymouth, NH, 03264
  - code: University of Ottawa
    address: Geography, Environment and Geomatics, 550 Cumberland St., Ottawa, ON, Canada, K1N 6N5
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4--1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1%. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: ../HGIS_paper.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-with-titles.csl
output: 
  rticles::elsevier_article:
    latex_engine: xelatex
    keep_tex: true
    citation_package: default
---

@@

@@

```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig1}Schematic of the open split inlet. The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
```

@@

# Acknowledgements

@@

# Supplementary Materials

>>>>

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface"
author:
  - name: Brett E. Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Joshua R. Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon L. Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation:  Woods Hole Oceanographic Institution,Plymouth State University
  - name: Sydney D. Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark L. Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark D. Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Environmental Science and Policy, 17 High St., Plymouth, NH, 03264
  - code: University of Ottawa
    address: Geography, Environment and Geomatics, 550 Cumberland St., Ottawa, ON, Canada, K1N 6N5
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4--1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1%. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: 12pt, review, 3p, authoryear
bibliography: ../HGIS_paper.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-with-titles.csl
output: 
  rticles::elsevier_article:
    latex_engine: xelatex
    keep_tex: true
    citation_package: default
header-includes: 
 - \usepackage[section]{placeins}
---

@@

\newpage

@@

@@

\section*{Acknowledgements}

@@

\newpage

\section*{Figures}

```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="60%", fig.cap = "\\label{fig1}Schematic of the open split inlet. The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
```

\newpage

\section*{Tables}

\newpage

\section*{Supplementary Materials}
